---
title: "Kandinsky hot/cold areas"
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: true
    toc_float: false
    self_contained: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(warning = F, collapse = TRUE, highlight = TRUE, comment = "#>")
```

```{=html}
<style type="text/css">
pre code {
max-width: 800px;  
white-space: pre-wrap;
}

pre code .comment {
color: #b0bec5;
}

#TOC {
position: fixed;
left: 20px;
top: 100px;
width: 350px;
background-color: #f8f9fa;
padding: 10px;
border-radius: 5px;
font-size: 14px;
max-height: 90vh;
overflow-y: auto;
}

body {
margin-left: 300px;
font-size: 12pt;

}
</style>
```

```{r setup,echo=F, message = F}
library(Kandinsky)
library(ggplot2)
library(dplyr)
library(patchwork)
library(UCell)
```

# Introduction

<hr />

The following code replicates the analysis presented in the Kandinsky paper to investigate spatial pattern of CD74 gene expression in a colorectal cancer (CRC) sample, using **CosMx** and **Visium** spatial transcriptomic data.

# CosMx

<hr />

Raw CosMx data and cell annotation files of CRC sample "CR48" were downloaded from [Zenodo](https://zenodo.org/records/10927005).

The dataset consists of 93 fields of view (FOVs) from a **human CRC FFPE tissue**, profiled using the *CosMx Universal Cell Characterization RNA Panel*.

Cell type annotation for all single cells in the dataset was generated with `InSituType` as part of an independent [study](https://www.cell.com/cancer-cell/fulltext/S1535-6108(24)00483-5).

<br />

## 1) Loading data

<hr />

In order to work with CosMx data using **Kandinsky**, we first need to set the **path to the folder** containing raw input data:


```{r set-cosmx-path, echo = F}
cosmx_path = '/Users/cgb01/repo/Kandinsky/data/CosMx/'
```

```{r, eval = F}
cosmx_path = 'Kandinsky/data/CosMx/'
```

```{r list-file}
list.files(cosmx_path)
```

Now we can import and load CosMx data into a `Seurat object` using the Kandinsky `prepare_cosmx_seurat()` function, by specifying:

-   the `path` to the input folder

-   the `dataset.id` which will be used to name the resulting Seurat object

Additional parameters include:

-   `pattern`: to selectively import only files whose filenames contain a specific string, useful when the folder includes input files from multiple datasets

-   `fovs`: to load only the cells located within a specified list of FOVs. To load the full dataset, you can set `fovs = NULL` (default)

```{r load-cosmx-data,fig.dim=c(4,4)}
cosmx = prepare_cosmx_seurat(path = cosmx_path, dataset.id = 'CR48', pattern = 'CR48')
```

<br />


## 2) Loading cell type annotation

<hr />

We can load the annotation file and store the cell type information within the `Seurat object` metadata.

```{r load-cell-type-annotation,fig.dim=c(12,9)}
# Add cell type annotation to CosMx metadata
anno_file = list.files(cosmx_path, pattern = 'CosMx_S0_processed_metadata_plus_cell_anno', full.names = T)
anno = readRDS(paste0(cosmx_path,anno_file))
head(anno)
anno = anno[anno$sample == 'CR48',]
anno = anno[,c('cell_ID','final_anno')]

anno = sf::st_drop_geometry(anno)
anno$cell_ID = gsub("c_1_","",anno$cell_ID)
rownames(anno) = anno$cell_ID
common = intersect(colnames(cosmx),rownames(anno))
cosmx = cosmx[,rownames(anno)]

cosmx = update_kandinsky(cosmx)
anno = anno[colnames(cosmx),]
cosmx$final_anno = as.character(anno$final_anno)

message('CosMx dataset composed of ',nrow(anno),' cells')

```

We can now visualize our dataset using the Kandinsky `KanPlot()` function, with cells coloured according to their annotated cell types:

```{r plot-cell-types,fig.dim=c(13,9)}
KanPlot(cosmx,feature = 'final_anno') + 
  scale_fill_manual(values = c('Epi' = 'green', 'FibroEndoMuscle' = 'magenta', 'Myeloid' = 'cyan', 'T/NK' = 'red', 'Plasma/B' = 'yellow'))
```

<br />

## 3) c-NB identification

<hr />

Now that the new `Seurat object` is ready, we use **Kandinsky** to identify **cell neighborhoods (c-NBs)** based on cell-to-cell relationships using the `kandinsky_init()` function.

Since we are working with CosMx data, we set the parameter `tech = 'cosmx'` to ensure Kandinsky correctly interprets the CosMx data format.

For neighborhood identification, we use the **membrane distance method** (`nb.method = 'M'`) and set the distance threshold to 30um (`d.max = 30`).

```{r init-kandinsky,fig.dim=c(4,4)}
cosmx = kandinsky_init(cosmx, tech = 'cosmx', nb.method = 'M', d.max = 30)
```

<br />

We can assess the **size of the identified c-NBs**, expressed as number of cells found within each c-NB, using the Kandinsky `nbSizePlot()` function:

```{r nbSize}
nbSizePlot(cosmx) + theme(aspect.ratio = 2)
```

Before continuing with the downstream analysis, we need to perform gene expression normalization using Seurat function `NormalizeData()`:

```{r normalization}
# Normalize counts
cosmx = NormalizeData(cosmx)
```

<br />

## 4) Hot/cold areas

<hr />

We can visualize single cell CD74 expression levels across different FOVs:

```{r cd74-hot-cold-areas,fig.dim=c(12,9)}
KanPlot(cosmx,'CD74', palette_cont = 'viridis')
```

We can define CD74 hot and cold areas according Getis-Ord Gi statistics calculated using Kandinsky function `hotspot_analysis()`.

In particular, we can specify as parameter:

-   `feature`: the name of the feature to use to compute Getis-Ord Gi\* statistics
-   `sim`: number of Monte Carlo simulations to be run for estimating Local Gi\* coefficients significance
-   `padj.thresh`: numeric value indicating the significance threshold to be applied to the adjusted p-values resulting from Monte Carlo simulations

```{r hotspot}
cosmx = hotspot_analysis(cosmx, feature = 'CD74', layer = 'data', sim=999, padj.thresh = 0.05)
```


We can now visualize  cells coloured according to whether they are assigned to hot or cold areas based on CD74 expression


```{r hotspot-plot}
KanPlot(cosmx, feature = 'CD74_clust', palette_discrete = c('dodgerblue3','firebrick3','#D9D9D9'))
```



After selecting a subset of representative FOVs, we can now visualize the same subset of cells coloured according to:

-   their cell type of origin
-   CD74 expression
-   whether they are assigned to hot or cold areas based on CD74 expression

```{r visualize-representative-fovs,fig.dim=c(12,9)}
fovs = c(154,157:159)

(KanPlot(cosmx, feature = 'final_anno', fovs = fovs
         , show_border = T, border.size = 0.1, border.col = '#808285') +
    scale_fill_manual(values = c('Epi' = 'green', 'FibroEndoMuscle' = 'magenta', 'Myeloid' = 'cyan', 'T/NK' = 'red', 'Plasma/B' = 'yellow'))) +
  
  KanPlot(cosmx, feature = 'CD74', fovs = fovs
          , palette_cont = 'viridis', show_border = T, border.size = 0.1, border.col = '#808285') +
  
  KanPlot(cosmx,feature = 'CD74_clust', fovs = fovs
          , palette_discrete = c('#1A6AA0','#A94349','#D9D9D9'), show_border = T, border.size = 0.1, border.col = '#808285')
```

We can test significant enrichment of annotated cell types within CD74 hot and cold areas using one-tailed Fisher’s exact test:

```{r cd74-celltype-overlap,fig.dim=c(5,9)}
cd74 = model.matrix(data = cosmx@meta.data, ~ 0 + CD74_clust) %>% as.data.frame()
colnames(cd74) = c('Cold','Hot','NS')

immune = model.matrix(data = cosmx@meta.data, ~0 + final_anno) %>% as.data.frame()
colnames(immune) = c('Epi','FibroEndoMuscle','Myeloid','Plasma/B','TNK')

immune  = immune[,c('Epi','Myeloid','TNK')]

all_res = lapply(colnames(immune),function(x){
  fisher_hot = fisher.test(table(cd74$Hot, immune[[x]]), alternative = 'greater')$p.value
  fisher_cold = fisher.test(table(cd74$Cold, immune[[x]]), alternative = 'greater')$p.value
  return(data.frame(cell_type = x
                    ,hot_pval = fisher_hot
                    ,cold_pval = fisher_cold))
})

all_res = purrr::reduce(all_res, rbind)
all_res$FDR = p.adjust(all_res$hot_pval, method = 'BH')


cosmx_hot = list()
for(c in c('Epi','Myeloid','TNK')){
  fisher_hot = all_res[all_res$cell_type == c,]$FDR
  
  if(fisher_hot < 2.2*(10^-16)){
    fisher_hot = expression('Fisher FDR < 2.2x10'^-16)
  }else{
    fisher_hot = paste0('Fisher FDR = ', format(fisher_hot, scientific = T, digits = 3))
  }
  
  hot_props = table(cd74$Hot,immune[[c]]) %>% as.data.frame() %>% 
    group_by(Var1) %>% 
    mutate(prop = Freq/sum(Freq)) %>% ungroup() %>% 
    filter(Var2 == 1)
  
  cosmx_hot[[c]] = ggplot(hot_props, aes(x = Var1, y = prop))+
    theme_classic() + geom_bar(stat = 'identity', fill = ifelse(
      c == 'Epi','green',ifelse(c == 'Myeloid','cyan','red')
    ), color = 'black') +
    scale_x_discrete(breaks = c('1','0'), limits = c('1','0'), labels = c('Hot','Rest')) +
    labs(x = 'CD74 areas', y = paste0('Area (%)')) +
    scale_y_continuous(limits = c(0, (max(hot_props$prop)*1.1)), breaks = c(0,max(hot_props$prop)), labels = round(c(0,100*(max(hot_props$prop))), 1))+
    annotate('text', x = 1.5, y = max(hot_props$prop)*1.1, label = fisher_hot) +
    guides(y = guide_axis(cap = "both"), x = guide_axis(cap = "both")) + 
    theme(axis.text = element_text(color='black'), aspect.ratio = 1)
}

wrap_plots(cosmx_hot[c('Myeloid','TNK','Epi')], ncol = 1)
```

The results from Fisher's tests indicate an enrichment of both myeloid and T/NK cells within CD74 hot areas compared to other regions of the sample.

Although CRC epithelial cells are not enriched in CD74 hot areas, findings from [Acha-Sagredo et al](https://www.cell.com/cancer-cell/fulltext/S1535-6108(24)00483-5) suggest that CRC cells can be induced to overexpress CD74 as a result of proximity to T/NK cell.

We can corroborate this observation by dividing our CRC cells into two groups, based on whether or not at least one T/NK cells is present within their c-NBs.\
We can then compare CD74 expression between these two groups of CRC cells via Wilcoxon Rank test.

```{r cd74-epi-expression,fig.dim=c(9,6)}
cosmx = nnMat(cosmx, label = 'final_anno')
cosmx = nn_query(cosmx
                 , query = "`T/NK` >0, final_anno == 'Epi'"
                 , anno_name = 'Epi_type'
                 , diffexpr = F
                 , label = 'final_anno'
                 , which = 'Epi')

cosmx$CD74_expr = FetchData(cosmx, 'CD74', clean = 'none')[,1]

cosmx_cd74_data = cosmx@meta.data[cosmx@meta.data$final_anno == 'Epi',c('Epi_type','CD74_expr')]

pval = wilcox.test(CD74_expr ~ Epi_type, data = cosmx_cd74_data)$p.value

if(pval < 2.2e-16){
  pval = expression('Wilcoxon p-value < 2.2x10'^-16)
}else{
  pval = format(pval, scientific = T, digits = 3)
}


n_epi1 = nrow(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'N',])
n_epi2 = nrow(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'Y',])

dens1 = density(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'N',]$CD74_expr)$y
dens2 = density(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'Y',]$CD74_expr)$y
max_dens = max(c(dens1,dens2))
min_dens = min(c(dens1,dens2))
median1 = median(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'N',]$CD74_expr)
median2 = median(cosmx_cd74_data[cosmx_cd74_data$Epi_type == 'Y',]$CD74_expr)

ggplot(cosmx_cd74_data, aes(x = CD74_expr, fill = Epi_type)) +
  theme_classic() + geom_density(alpha = 0.4,color=NA)+
  scale_x_continuous(breaks = c(min(cosmx_cd74_data$CD74_expr), median1, median2, max(cosmx_cd74_data$CD74_expr))
                     , labels = round(c(min(cosmx_cd74_data$CD74_expr), median1, median2, max(cosmx_cd74_data$CD74_expr)), 1)) +
  scale_y_continuous(breaks = c(min_dens, max_dens)
                     , labels = round(c(min_dens, max_dens),2)) +
  guides(y = guide_axis(cap = "both"), x = guide_axis(cap = "both")) +
  labs(x = 'CD74 normalized expression', y = 'Density', fill = 'Epi type') +
  annotate('text', x = median(cosmx_cd74_data$CD74_expr), y = max_dens, label = pval) +
  scale_fill_manual(values = c('#2E8B57','#80E580')
                    , labels = c(paste0('Rest of CRC cells (', n_epi1, ')'), paste0('CRC cells near T/NK (', n_epi2, ')'))) +
  geom_vline(xintercept = median1, linetype = 'dashed', color = '#2E8B57') +
  geom_vline(xintercept = median2, linetype = 'dashed', color = '#80E580') + 
  coord_fixed(10)
```

Indeed, CRC cells with at least one T/NK cell within their c-NBs show a significantly higher CD74 expression compared to the rest of CRC cells.

<br />

# Visium

<hr />

The **Visium dataset** analysed in this vignette has been downloaded from [Zenodo]().

It consists of a distinct tissue regions obtained from the same **human CRC FFPE sample** previously profiled with the *CosMx Universal Cell Characterization RNA Panel*.

The *Visium Human Transcriptome Probe kit* (v1) was used to capture RNA molecules from the FFPE sample after tissue permeabilization, enabling gene expression quantification for \~18,000 human genes.

After sequencing, Visium FASTQ files were processed with spaceranger (v2.0). Loupe Browser (v6.2) was used to visualize Visium spot-level gene expression data and the associated H&E image, to check for the presence of "in-tissue" spots located in empty regions of the tissue slide.

A separate spot annotation file `A1_CR48_TissueType_Anno.csv` was generated.

## 1) Loading data

<hr />

We first need to specify the **path to the folder** containing raw Visium input data.\
Next, we import Visium data into R as a `Seurat object` using the `Load10X_Spatial()` function.

Once the data is loaded, we perform initial quality checks. Specifically:

-   Visium spots mapped to empty tissue regions are removed based on the `A1_CR48_TissueType_Anno.csv` annotation file
-   Spots with fewer than 500 transcripts (UMI) and fewer than 300 unique genes detected are discarded
-   Genes detected in fewer than 10 spots (after filtering out low-quality spots) are removed


```{r, echo = FALSE}
visium_path = '/Users/cgb01/repo/Kandinsky/data/Visium/'
```

```{r, eval = FALSE}
visium_path = 'Kandinsky/data/Visium/'
```

```{r load-and-preprocess-visium}
visium = Load10X_Spatial(
  , data.dir = visium_path
  , filename = "filtered_feature_bc_matrix.h5"
  , assay = "Spatial"
  , slice = 'CR48'
  , filter.matrix = TRUE
)   
genes = rownames(visium)

message("Total spots: ",ncol(visium))

# Load spot annotation file (used to remove 'empty' spots)
anno_spots = read.delim(paste0(visium_path,'A1_CR48_TissueType_Anno.csv'), sep = ',')
rownames(anno_spots) = anno_spots$Barcode

# Remove spots without tissue annotation (corresponding to empty regions)
anno_spots = anno_spots[which(anno_spots$Location != ""),]


# Remove spots corresponding to empty regions (if any)
if(ncol(visium) > nrow(anno_spots)){
  message("Removing ", ncol(visium) - nrow(anno_spots), " empty spots")
  visium = visium[,which(colnames(visium) %in% rownames(anno_spots))]
} else {
  message("No empty spots detected")
}

anno_spots = anno_spots[colnames(visium),]
visium = AddMetaData(visium, anno_spots$Location, col.name = "Location")


# Remove low-quality spots
visium = subset(x = visium, subset = nCount_Spatial >= 500)
message(ncol(visium)," spots with at least 500 UMIs")

visium = subset(x = visium, subset = nFeature_Spatial >= 300)
message(ncol(visium)," spots with at least 300 genes detected")
cell_id = colnames(visium)

keep_genes = which(rowSums(LayerData(visium,'counts') > 0) >= 10)
visium = subset(visium, features = keep_genes)
dim(visium)
```

<br />

## 2) s-NB identification

<hr />

Using Kandinsky `kandinsky_init()` function, we define spot neighbourhoods (s-NBs) based on the **Queen contiguity method** (`nb.method = ‘Q’`), selecting the first layer of neighboring spots around each spot

We then use the Kandinsky `nb_expand()` function to include the **second closest layer** of spots in each s-NB.

```{r kandinsky-init, fig.dim=c(4,8)}
visium = kandinsky_init(visium
                        , tech ='visium'
                        , nb.method = 'Q'
                        , res = 'high'
                        , img = paste0(visium_path,'spatial/tissue_hires_image.png')
)

# Expand neighbourhood definition by including the second ring of surrounding spots
visium = nb_expand(visium, maxorder = 2)
```

Moreover, we can assess the **size of identified s-NBs**, expressed as number of cells found within each s-NB, using the Kandinsky `nbSizePlot()` function:

```{r}
nbSizePlot(visium) + theme(aspect.ratio = 2)
```

The H&E image associated with the Visium sample is automatically stored in the Kandinsky slot `img`.\
For visualization purposes, the user can create a **mask** of the H&E tissue slide, retaining only the image area covered by tissue, using the Kandinsky `he_mask()` function.

```{r mask-he,fig.dim=c(6,6)}
# Create a mask image starting from visium H&E staining image
visium = he_mask(visium, stretch = T, sd_thresh = 7)

SlidePlot(visium, img_type = 'img') + coord_fixed(1) +
SlidePlot(visium, img_type = 'mask') + coord_fixed(1)
```

Before continuing with the downstream analysis, we need to perform gene expression normalization using Seurat function `NormalizeData()`:

```{r normalize-cd74-expression,fig.dim=c(10,10)}
visium = NormalizeData(visium)
```

<br />

## 3) Hot/cold areas

<hr />

We can visualize CD74 expression levels across spots:

```{r}
KanPlot(visium, feature = 'CD74', palette_cont = 'viridis')
```

Now, we can define CD74 hot and cold areas according to Getis-Ord Gi statistics calculated using Kandinsky `hotspot_analysis()` function.

```{r define-cd74-hot-cold-areas}
# Define hot/cold areas of CD74 expression (default values used are the normalized counts stored in the 'data' layer)
visium = hotspot_analysis(visium, feature='CD74', sim = 999, padj.thresh = 0.05)

KanPlot(visium, feature = 'CD74_clust', palette_discrete = c('#1A6AA0','#A94349','#D9D9D9'))
```

Since Visium spots cover tissue regions that can generally contain multiple individual cells, we can use **cell type gene expression signatures** to estimate the abundance of different cell populations within each spot.

We load the list of **CRC immune-related signatures** described in [Acha-Sagredo et al](https://www.cell.com/cancer-cell/fulltext/S1535-6108(24)00483-5). Based on our analysis of the CosMx CRC dataset, we revealed an association between CD74 hot areas and myeloid and T/NK cells.

Now, we are interested in defining the **relationship** between **CD74 expression** and specific subpopulations of tumor-associated **macrophages** and **T/NK cells**.

```{r, echo = FALSE}
signatures = readRDS('/Users/cgb01/repo/Kandinsky/data/Visium/CRC_LCM_Ext_sigs.rds')
```

```{r load-signatures, eval = FALSE}
signatures = readRDS('Kandinsky/data/Visium/CRC_LCM_Ext_sigs.rds')
```

We can calculate gene expression scores across spots for each immune-related signature using `UCell`.\
Next, we define **hot and cold areas** for each immune signature using Kandinsky `hotspot_analysis()` function, considering the signature scores calculated with `UCell` across all spots .\

```{r immune-hot-cold-areas-detection}
# Test signatures related to TAMs/TNK subtypes or IFN production
signatures = signatures[c(4,5,6,10,12,13,14)]

# Calculate signature scores within each spot using UCell
visium = UCell::AddModuleScore_UCell(visium, features = signatures, maxRank = 5000)

visium = visium %>% 
  FindVariableFeatures(., nfeatures = 5000, selection.method = "vst") %>% 
  ScaleData() %>%
  RunPCA(assay = "Spatial", verbose = FALSE,npcs = 50) %>%
  FindNeighbors(reduction = "pca", dims = 1:30)

# Smooth UCell scores across spots according to their 20 nearest neighbour in the PCA space
visium = UCell::SmoothKNN(visium, signature.names = paste0(names(signatures),'_UCell'), k = 20)

# Immune sig hot/cold areas
for(s in paste0(names(signatures),'_UCell_kNN')){
  visium = hotspot_analysis(visium, feature = s, sim = 999, padj.thres = 0.05)
}

# Clean column names for hot/cold areas annotation
colnames(visium@meta.data) = gsub('_UCell_kNN' ,'', colnames(visium@meta.data))
```

<br />

When hot and cold areas are identified for more than one variable within the same dataset, the Kandinsky `hotspot_overlap` function can be used to create a new annotation that reflects the **overlap** between hot and cold areas for **two independent variables**, specified using `feat.1` and `feat.2` parameters.

```{r hotspot-overlap,fig.dim=c(6,6)}
KanPlot(visium, feature = 'M1_16_clust', palette_discrete = c('#1A6AA0','#A94349','#D9D9D9'))

visium = hotspot_overlap(visium, feat.1 = 'CD74', feat.2 = 'M1_16')

KanPlot(visium, feature = 'CD74_M1_16_clust', palette_discrete = c('dodgerblue4','skyblue1','purple','firebrick4','firebrick1','gray85'), border.col = 'gray95')
KanPlot(visium, feature = 'CD74_M1_16_clust', palette_discrete = c('gray85','skyblue1','gray85','gray85','firebrick1','gray85'), border.col = 'gray95')

```

We can **test** the association between CD74 hot and cold areas and each TAM or T/NK related signature using a one-tailed Fisher’s Exact test.\
Specifically, we assess the overlap between CD74 hot/cold areas and hot/cold areas for each signature, respectively:

```{r hot-cold-areas-enrich}
res = list()

for(s in paste0(names(signatures),'_clust')){
  cd74 = model.matrix(data = visium@meta.data, ~ 0 + CD74_clust) %>% as.data.frame()
  colnames(cd74) = c('Cold','Hot','NS')
  
  immune = model.matrix(data = visium@meta.data, formula(paste0("~0+", s))) %>% as.data.frame()
  colnames(immune) = c('Cold','Hot','NS')
  
  fisher_hot = fisher.test(table(cd74$Hot, immune$Hot), alternative = 'greater')$p.value
  fisher_cold = fisher.test(table(cd74$Cold, immune$Cold), alternative = 'greater')$p.value
  tot_res = data.frame(signature = gsub("_clust","", s),
                       hot_hot_pval = fisher_hot,
                       cold_cold_pval = fisher_cold)
  res[[ gsub("_clust","",s)]] = tot_res
}

res = purrr::reduce(res,rbind)

res$hot_FDR = p.adjust(res$hot_hot_pval, method = 'BH')
res$cold_FDR = p.adjust(res$cold_cold_pval, method = 'BH')
rownames(res) = res$signature


summary_res = res %>% dplyr::select(signature, hot_FDR, cold_FDR)
summary_res = reshape2::melt(summary_res)
```

We can now visualize the association between CD74 expression levels and each TAM or T/NK related signature by

```{r plot-hot-cold-enrich,fig.dim=c(12,7)}
plots_hot = list()

for(s in paste0(names(signatures),'_clust')){
  print(s)
  cd74 = model.matrix(data = visium@meta.data, ~ 0 + CD74_clust) %>% as.data.frame()
  colnames(cd74) = c('Cold','Hot','NS')
  
  immune = model.matrix(data = visium@meta.data, formula(paste0("~0+",s))) %>% as.data.frame()
  colnames(immune) = c('Cold','Hot','NS')
  
  fisher_hot = summary_res[summary_res$signature == gsub('_clust','',s) & summary_res$variable == 'hot_FDR',]$value
  
  if(fisher_hot < 2.2e-16){
    fisher_hot = expression('Fisher FDR < 2.2x10'^-16)
  }else{
    fisher_hot = paste0('Fisher FDR = ',format(fisher_hot, scientific = TRUE, digits = 3))
  }
  
  hot_props = table(cd74$Hot,immune$Hot) %>% as.data.frame() %>% 
    group_by(Var1) %>% 
    mutate(prop = Freq/sum(Freq)) %>% ungroup() %>% 
    filter(Var2 == 1)
  
  plots_hot[[s]] = ggplot(hot_props,aes(x = Var1, y = prop)) + 
    theme_classic() + geom_bar(stat = 'identity', fill = '#A94349', color = 'black') +
    scale_x_discrete(breaks = c('1','0'), limits = c('1','0'), labels = c('Hot','Rest')) +
    labs(x = 'CD74 area', y = paste0('Hot for ',gsub('_clust','',s),' (%)'), fill = paste0(gsub('_clust','',s),' area')) +
    scale_y_continuous(limits = c(0,max(hot_props$prop)+0.05)
                       , breaks = c(0,max(hot_props$prop))
                       ,labels = round(100*c(0,max(hot_props$prop)),1)) +
    annotate('text', x = 1.5, y = max(hot_props$prop) + 0.05, label = fisher_hot)+
    guides(y = guide_axis(cap = "both"), x = guide_axis(cap = "both")) + 
    theme(axis.text = element_text(color = 'black'))
}

wrap_plots(plots_hot, ncol = 4)

plots_cold = list()
for(s in paste0(names(signatures),'_clust')){
  print(s)
  cd74 = model.matrix(data = visium@meta.data, ~0 + CD74_clust) %>% as.data.frame()
  colnames(cd74) = c('Cold','Hot','NS')
  
  immune = model.matrix(data = visium@meta.data, formula(paste0("~0+",s))) %>% as.data.frame()
  colnames(immune) = c('Cold','Hot','NS')
  
  fisher_cold = summary_res[summary_res$signature == gsub('_clust','',s) & summary_res$variable == 'cold_FDR',]$value
  if(fisher_cold < 2.2e-16){
    fisher_cold = expression('Fisher FDR < 2.2x10'^-16)
  }else{
    fisher_cold = paste0('Fisher FDR = ',format(fisher_cold, scientific = T, digits = 3))
  }
  
  cold_props = table(cd74$Cold, immune$Cold) %>% as.data.frame() %>% 
    group_by(Var1) %>% 
    mutate(prop = Freq/sum(Freq)) %>% ungroup() %>% 
    filter(Var2 == 1)
  
  plots_cold[[s]] = ggplot(cold_props,aes(x = Var1, y = prop)) + 
    theme_classic() + geom_bar(stat = 'identity', fill = '#1A6AA0', color = 'black')+
    scale_x_discrete(breaks = c('1','0')
                     , limits = c('1','0')
                     , labels = c('Cold','Rest')) +
    labs(x = 'CD74 area', y = paste0('Cold for ',gsub('_clust','',s),' (%)'), fill = paste0(gsub('_clust','',s),' area')) +
    scale_y_continuous(limits = c(0, max(cold_props$prop)+0.05)
                       , breaks = c(0, max(cold_props$prop))
                       , labels = round(100*c(0, max(cold_props$prop)),1)) +
    annotate('text', x = 1.5, y = max(cold_props$prop) + 0.05, label = fisher_cold) +
    guides(y = guide_axis(cap = "both"), x = guide_axis(cap = "both")) + 
    theme(axis.text = element_text(color = 'black'))
}
wrap_plots(plots_cold, ncol = 4)
```
